Follow-up Samples

### 2016-12-07 submission/manuscript_outline_20161024.Rmd
results update of text

***Bacterial Community and Fit Changes before and after Treatment***
Based on thetayc distance metrics comparing the initial to the follow up samples there was no difference between the adenoma and carcinoma group (P-value = `r format(change_summary['thetayc', 'Pvalue'], scientific = FALSE, digits = 3)`) [Figure 1a]. There was a difference in FIT between initial and follow up samples with the carcinoma group having a significant decrease in FIT versus the adenoma group (P-value = `r format(change_summary['fit', 'Pvalue'], scientific = TRUE, digits = 3)`) [Figure 1b].  Although the thetayc distance metric change was similar between adenoma and carcinoma the directionality of the change was significant in the cancer group between initial and follow up (P-value = `r format(beta_summary['Carcinoma', 'PERMANOVA'], scientific = FALSE, digits = 3)`) but not for the adenoma group (P-value = `r format(beta_summary['Adenoma', 'PERMANOVA'], scientific = FALSE, digits = 3)`) [Figure 2]. When all follow up samples were compared to each other there was no significant overall difference between them (P-value = `r format(beta_summary['All', 'PERMANOVA'], scientific = FALSE, digits = 3)`).  There was no significant difference between initial and follow up samples for observed OTUs, Shannon diversity, and evenness after correction for multiple comparisons [Table S1].  Time of follow up sample from initial sampling did not have an effect on our prediction data set (uncorrected P-value = `r format(time_pvalues["thetayc_change", "uncorrected_p_value"], digits = 3)`).

***Outcome of Model Training***
The range of the AUC for model training ranged from a minimum of `r format(model_test_AUC_summary["worse", "ROC_cv"], digits = 3)` to a maximum of `r format(model_test_AUC_summary["best", "ROC_cv"], digits = 3)`.  To be conservative the model chosen for prediction of disease in the follow up samples had an AUC in the middle of all the `r length(rownames(model_train_summary))` runs which was `r format(model_test_AUC_summary["middle", "ROC_cv"], 
digits = 3)`.  Interestingly, the worst AUC model from training performed the best on it's respective 80/20 split test data [Figure 3].  In fact the 80/20 test performance showed that the AUC for the middle model chosen was the most stable (best training model test set AUC = `r format(model_test_AUC_summary["best", "AUC"], digits = 3)`, middle training model test set AUC = `r format(model_test_AUC_summary["middle", "AUC"], digits = 3)`, worse training model test set AUC = `r format(model_test_AUC_summary["worse", "AUC"], digits = 3)`).  That is to say it had the smallest change in AUC in comparison to the other minimum and maximum AUC trained models.  There was no significant difference between the AUC of the best and middle training models (P-value = `r format(roc_pvalues["best", "middle"], digits = 3)`).  There was also no difference in the middle model versus worse (P-value = `r format(roc_pvalues["middle", "worse"], digits = 3)`) or full data model (P-value = `r format(roc_pvalues["middle", "full"], digits = 3)`).  The two comparisons with a significant difference were between the worse training model and the best training model (P-value = `r format(roc_pvalues["best", "worse"], scientific = TRUE, digits = 3)`) and the worse training model and full data model (P-value = `r format(roc_pvalues["worse", "full"], scientific = TRUE, digits = 3)`).

@@

***Surgical Removal of an Adenoma or Carcinoma Results in a Decrease in Positive Probability Prediction***
A total of `r length(rownames(good_metaf)) - length(rownames(filter(good_metaf, !is.na(fit_followUp))))` sample was omitted from the original `r length(rownames(good_metaf))` sample set since it was missing a complete set of follow up data.  This left a total of `r sum(initial_cm$ref_yes)` samples for test predictions.  After multiple comparison correction there was a significant overall decrease in positive probability of a carcinoma and adenoma (P-value = `r format(IF_model_wilcox_paired_summary["lesion", "BH_correction"], scientific = TRUE, digits = 3)`) [Figure 4].  This decrease was significant for both adenoma (P-value = `r format(IF_model_wilcox_paired_summary["all_adenoma", "BH_correction"], scientific = TRUE, digits = 3)` [Figure 4a] and carcinoma (P-value = `r format(IF_model_wilcox_paired_summary["carcinoma_only", "BH_correction"], scientific = TRUE, digits = 3)`) [Figure 4b] alone.  This also held specifically for those with screen relative neoplasias (SRN) (P-value = `r format(IF_model_wilcox_paired_summary["SRN_only", "BH_correction"], scientific = TRUE, digits = 3)`).  A total of `r initial_cm["pred_yes", "ref_yes"]` or `r format(confusion_summary["Accuracy", "initial"]*100, digits = 3)`% of all samples were correctly predicted to have an adenoma or carcinoma.  Although there was a decrease in positive probability only `r followup_cm["pred_no", "ref_no"]` of the total `r sum(initial_cm$ref_yes)` individuals were classified as adenoma or carcinoma free on follow up (successful classification of `r format(confusion_summary["Accuracy", "followup"]*100, digits = 3)`%). There was no significant difference between the predictions and actual diagnosis for the initial samples in the 67-sample cohort test set.  However, the predictions were significantly discordent with the diagnosis for the follow up samples (P-value = `r format(confusion_summary["McnemarPValue", "followup"], scientific = TRUE, digits = 3)`).  Although there were discordent results the respective sensitivity for the initial group was `r format(confusion_summary["Sensitivity", "initial"]*100, digits = 3)`% and for follow up was `r format(confusion_summary["Sensitivity", "followup"]*100, digits = 3)`%, respectively.

>>>>

***Bacterial Community and Fit Changes before and after Treatment***
Based on thetayc distance metrics, comparing the initial to the follow up samples, there was no difference between the adenoma and carcinoma groups (P-value = `r format(change_summary['thetayc', 'Pvalue'], scientific = FALSE, digits = 3)`) [Figure 1a]. There was a difference in FIT between initial and follow up samples with the carcinoma group having a significant decrease in FIT versus the adenoma group (P-value = `r format(change_summary['fit', 'Pvalue'], scientific = TRUE, digits = 3)`) [Figure 1b].  Although the thetayc distance metric change was similar between adenoma and carcinoma the directionality of the change was significant in the cancer group between initial and follow up (P-value = `r format(beta_summary['Carcinoma', 'PERMANOVA'], scientific = FALSE, digits = 3)`) but not for the adenoma group (P-value = `r format(beta_summary['Adenoma', 'PERMANOVA'], scientific = FALSE, digits = 3)`).  This change can be visualized using an NMDS [Figure 2]. When all follow up samples were compared to each other there was no significant overall difference between them (P-value = `r format(beta_summary['All', 'PERMANOVA'], scientific = FALSE, digits = 3)`).  There was no significant difference between initial and follow up samples for observed OTUs, Shannon diversity, or evenness after correction for multiple comparisons [Table S1].  Time of follow up sample from initial sampling, did not have a significant difference between adenoma and carcinoma (uncorrected P-value = `r format(time_pvalues["thetayc_change", "uncorrected_p_value"], digits = 3)`).

***Outcome of Model Training***
The range of the AUC for model training ranged from a minimum of `r format(model_test_AUC_summary["worse", "ROC_cv"], digits = 3)` to a maximum of `r format(model_test_AUC_summary["best", "ROC_cv"], digits = 3)` with the middle of all `r length(rownames(model_train_summary))` runs having an AUC of `r format(model_test_AUC_summary["middle", "ROC_cv"], 
digits = 3)`.  Interestingly, the worst AUC model from training performed the best on it's respective 80/20 split test data [Figure 3].  In fact the 80/20 test performance showed that the AUC for the middle model chosen was the most stable (best training model test set AUC = `r format(model_test_AUC_summary["best", "AUC"], digits = 3)`, middle training model test set AUC = `r format(model_test_AUC_summary["middle", "AUC"], digits = 3)`, worse training model test set AUC = `r format(model_test_AUC_summary["worse", "AUC"], digits = 3)`).  That is to say it had the smallest change in AUC in comparison to the minimum and maximum AUC trained models.  The middle model was close to the full training data AUC `r format(model_test_AUC_summary["full", "AUC"], digits = 3)`. There was no significant difference between the AUC of the best and middle training models (P-value = `r format(roc_pvalues["best", "middle"], digits = 3)`).  There was also no difference in the middle model versus worse (P-value = `r format(roc_pvalues["middle", "worse"], digits = 3)`) or full data model (P-value = `r format(roc_pvalues["middle", "full"], digits = 3)`).  The two comparisons with a significant difference were between the worse training model and the best training model (P-value = `r format(roc_pvalues["best", "worse"], scientific = TRUE, digits = 3)`) and the worse training model and full training model (P-value = `r format(roc_pvalues["worse", "full"], scientific = TRUE, digits = 3)`).

@@

***Surgical Removal of an Adenoma or Carcinoma Results in a Decrease in Positive Probability Prediction***
A total of `r length(rownames(good_metaf)) - length(rownames(filter(good_metaf, !is.na(fit_followUp))))` sample was omitted from the original `r length(rownames(good_metaf))` test sample set since it was missing a complete set of follow up data.  This left a total of `r sum(initial_cm$ref_yes)` samples for test predictions.  After multiple comparison correction there was a significant overall decrease in positive probability of a carcinoma and adenoma (P-value = `r format(IF_model_wilcox_paired_summary["lesion", "BH_correction"], scientific = TRUE, digits = 3)`) [Figure 4].  This decrease was significant for both adenoma (P-value = `r format(IF_model_wilcox_paired_summary["all_adenoma", "BH_correction"], scientific = TRUE, digits = 3)` [Figure 4a] and carcinoma (P-value = `r format(IF_model_wilcox_paired_summary["carcinoma_only", "BH_correction"], scientific = TRUE, digits = 3)`) [Figure 4b] alone.  This decrease in probability of lesion also held specifically for those with screen relative neoplasias (SRN) (P-value = `r format(IF_model_wilcox_paired_summary["SRN_only", "BH_correction"], scientific = TRUE, digits = 3)`).  A total of `r initial_cm["pred_yes", "ref_yes"]` or `r format(confusion_summary["Accuracy", "initial"]*100, digits = 3)`% of all samples were correctly predicted to have a lesion.  Although there was a decrease in positive probability only `r followup_cm["pred_no", "ref_no"]` of the total `r sum(initial_cm$ref_yes)` individuals were classified as adenoma or carcinoma free on follow up (successful classification of `r format(confusion_summary["Accuracy", "followup"]*100, digits = 3)`%). There was no significant difference between the predictions and actual diagnosis for the initial samples in the `r length(rownames(good_metaf))`-sample cohort test set.  However, the predictions were significantly discordent with the diagnosis for the follow up samples (P-value = `r format(confusion_summary["McnemarPValue", "followup"], scientific = TRUE, digits = 3)`).  Although there were discordent results the respective sensitivity for the initial group was `r format(confusion_summary["Sensitivity", "initial"]*100, digits = 3)`% and for follow up was `r format(confusion_summary["Sensitivity", "followup"]*100, digits = 3)`%, respectively.

### 2016-12-07 submission/manuscript_outline_20161024.tex
results update of text

\textbf{\emph{Bacterial Community and Fit Changes before and after
Treatment}} Based on thetayc distance metrics comparing the initial to
the follow up samples there was no difference between the adenoma and
carcinoma group (P-value = 0.697) {[}Figure 1a{]}. There was a
difference in FIT between initial and follow up samples with the
carcinoma group having a significant decrease in FIT versus the adenoma
group (P-value = 2.15e-05) {[}Figure 1b{]}. Although the thetayc
distance metric change was similar between adenoma and carcinoma the
directionality of the change was significant in the cancer group between
initial and follow up (P-value = 0.002) but not for the adenoma group
(P-value = 0.997) {[}Figure 2{]}. When all follow up samples were
compared to each other there was no significant overall difference
between them (P-value = 0.085). There was no significant difference
between initial and follow up samples for observed OTUs, Shannon
diversity, and evenness after correction for multiple comparisons
{[}Table S1{]}. Time of follow up sample from initial sampling did not
have an effect on our prediction data set (uncorrected P-value = 0.784).

\textbf{\emph{Outcome of Model Training}} The range of the AUC for model
training ranged from a minimum of 0.795 to a maximum of 0.854. To be
conservative the model chosen for prediction of disease in the follow up
samples had an AUC in the middle of all the 100 runs which was 0.819.
Interestingly, the worst AUC model from training performed the best on
it's respective 80/20 split test data {[}Figure 3{]}. In fact the 80/20
test performance showed that the AUC for the middle model chosen was the
most stable (best training model test set AUC = 0.682, middle training
model test set AUC = 0.815, worse training model test set AUC = 0.931).
That is to say it had the smallest change in AUC in comparison to the
other minimum and maximum AUC trained models. There was no significant
difference between the AUC of the best and middle training models
(P-value = 0.419). There was also no difference in the middle model
versus worse (P-value = 0.178) or full data model (P-value = 1). The two
comparisons with a significant difference were between the worse
training model and the best training model (P-value = 7.94e-04) and the
worse training model and full data model (P-value = 3.52e-03).

@@

\textbf{\emph{Surgical Removal of an Adenoma or Carcinoma Results in a
Decrease in Positive Probability Prediction}} A total of 1 sample was
omitted from the original 67 sample set since it was missing a complete
set of follow up data. This left a total of 66 samples for test
predictions. After multiple comparison correction there was a
significant overall decrease in positive probability of a carcinoma and
adenoma (P-value = 1.14e-11) {[}Figure 4{]}. This decrease was
significant for both adenoma (P-value = 7.95e-08 {[}Figure 4a{]} and
carcinoma (P-value = 7.95e-08) {[}Figure 4b{]} alone. This also held
specifically for those with screen relative neoplasias (SRN) (P-value =
5.35e-04). A total of 66 or 100\% of all samples were correctly
predicted to have an adenoma or carcinoma. Although there was a decrease
in positive probability only 24 of the total 66 individuals were
classified as adenoma or carcinoma free on follow up (successful
classification of 37.9\%). There was no significant difference between
the predictions and actual diagnosis for the initial samples in the
67-sample cohort test set. However, the predictions were significantly
discordent with the diagnosis for the follow up samples (P-value =
4.19e-10). Although there were discordent results the respective
sensitivity for the initial group was 100\% and for follow up was 100\%,
respectively.

>>>>

\textbf{\emph{Bacterial Community and Fit Changes before and after
Treatment}} Based on thetayc distance metrics, comparing the initial to
the follow up samples, there was no difference between the adenoma and
carcinoma groups (P-value = 0.697) {[}Figure 1a{]}. There was a
difference in FIT between initial and follow up samples with the
carcinoma group having a significant decrease in FIT versus the adenoma
group (P-value = 2.15e-05) {[}Figure 1b{]}. Although the thetayc
distance metric change was similar between adenoma and carcinoma the
directionality of the change was significant in the cancer group between
initial and follow up (P-value = 0.002) but not for the adenoma group
(P-value = 0.997). This change can be visualized using an NMDS {[}Figure
2{]}. When all follow up samples were compared to each other there was
no significant overall difference between them (P-value = 0.085). There
was no significant difference between initial and follow up samples for
observed OTUs, Shannon diversity, or evenness after correction for
multiple comparisons {[}Table S1{]}. Time of follow up sample from
initial sampling, did not have a significant difference between adenoma
and carcinoma (uncorrected P-value = 0.784).

\textbf{\emph{Outcome of Model Training}} The range of the AUC for model
training ranged from a minimum of 0.795 to a maximum of 0.854 with the
middle of all 100 runs having an AUC of 0.819. Interestingly, the worst
AUC model from training performed the best on it's respective 80/20
split test data {[}Figure 3{]}. In fact the 80/20 test performance
showed that the AUC for the middle model chosen was the most stable
(best training model test set AUC = 0.682, middle training model test
set AUC = 0.815, worse training model test set AUC = 0.931). That is to
say it had the smallest change in AUC in comparison to the minimum and
maximum AUC trained models. The middle model was close to the full
training data AUC 0.813. There was no significant difference between the
AUC of the best and middle training models (P-value = 0.419). There was
also no difference in the middle model versus worse (P-value = 0.178) or
full data model (P-value = 1). The two comparisons with a significant
difference were between the worse training model and the best training
model (P-value = 7.94e-04) and the worse training model and full
training model (P-value = 3.52e-03).

@@

\textbf{\emph{Surgical Removal of an Adenoma or Carcinoma Results in a
Decrease in Positive Probability Prediction}} A total of 1 sample was
omitted from the original 67 test sample set since it was missing a
complete set of follow up data. This left a total of 66 samples for test
predictions. After multiple comparison correction there was a
significant overall decrease in positive probability of a carcinoma and
adenoma (P-value = 1.14e-11) {[}Figure 4{]}. This decrease was
significant for both adenoma (P-value = 7.95e-08 {[}Figure 4a{]} and
carcinoma (P-value = 7.95e-08) {[}Figure 4b{]} alone. This decrease in
probability of lesion also held specifically for those with screen
relative neoplasias (SRN) (P-value = 5.35e-04). A total of 66 or 100\%
of all samples were correctly predicted to have a lesion. Although there
was a decrease in positive probability only 24 of the total 66
individuals were classified as adenoma or carcinoma free on follow up
(successful classification of 37.9\%). There was no significant
difference between the predictions and actual diagnosis for the initial
samples in the 67-sample cohort test set. However, the predictions were
significantly discordent with the diagnosis for the follow up samples
(P-value = 4.19e-10). Although there were discordent results the
respective sensitivity for the initial group was 100\% and for follow up
was 100\%, respectively.
